---
title: "COVID-19 Case Report"
output: html_document
---

## 0. Set up and Helper Functions

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readr)
library(xlsx)
library(caret)
library(car)
library(InformationValue)
library(caTools)
library(ggpubr)
library(stringr)
library(tidytext)
library(plotly)
library(tidyverse)

# helper functions
data_report_cont <- function(dataset) {
  feature <- colnames(dataset)
  count <- dim(dataset)[1]
  missing <- sapply(dataset, function(x) sum(is.na(x))/count)
  card <- sapply(dataset, function(x) length(unique(x[!is.na(x)])))
  summary_df <- t(sapply(dataset, function(x) summary(x[!is.na(x)])))
  std <- sapply(dataset, function(x) sd(x[!is.na(x)]))
  df <- tibble("Feature" = feature, "Count" = count, 
                   "Missing %" = missing, 
                   "Card." = card, "Min" = summary_df[,1], 
                   "1st Qrt." = summary_df[,2],
                   "Median" = summary_df[,3],
                   "Mean" = summary_df[,4],
                   "3rd Qrt." = summary_df[,5],
                   "Max" = summary_df[,6],
                   "Std. Dev." = std)
  return(df)
}

data_report_disc <- function(dataset) {
  feature <- colnames(dataset)
  count <- dim(dataset)[1]
  missing <- sapply(dataset, function(x) sum(is.na(x))/count)
  card <- sapply(dataset, function(x) length(unique(x[!is.na(x)])))
  mode <- sapply(dataset, function(x) names(sort(table(x), 
                                                 decreasing = T))[1])
  mode_freq <- sapply(dataset, function(x) sort(table(x), 
                                                decreasing = T)[1])
  mode_percent <- mode_freq/count
  second_mode <- sapply(dataset, function(x) names(sort(table(x), 
                                                        decreasing = T))[2])
  second_mode_freq <- sapply(dataset, function(x) sort(table(x), 
                                                       decreasing = T)[2])
  second_mode_percent <- second_mode_freq/count
  df <- tibble("Feature" = feature, "Count" = count, 
                   "Missing" = missing, "Card." = card,
                   "Mode" = mode, "Mode Freq." = mode_freq, 
                   "Mode %" = mode_percent,
                   "2nd Mode" = second_mode, "2nd Mode Freq." = second_mode_freq, 
                   "2nd Mode %" = second_mode_percent)
  return(df)
}


general_data_report <- function(dataset) {
  feature <- colnames(dataset)
  count <- dim(dataset)[1]
  missing <- sapply(dataset, function(x) sum(is.na(x))/count)
  card <- sapply(dataset, function(x) length(unique(x[!is.na(x)])))
  df <- tibble("Feature" = feature, "Count" = count, 
                   "Missing" = missing, "Cardinality" = card)
  return(df)
}


clean_txt <- function(txt){
    # lower case
    temp <- tolower(txt)
    # Remove everything that is not a number or letter 
    temp <- stringr::str_replace_all(temp,"[^a-zA-Z\\s]", " ")
    # Shrink down to just one white space
    temp <- stringr::str_replace_all(temp,"[\\s]+", " ")
    return(temp)
}


plot_multi_histogram <- function(df, feature, label_column) {
  plt <- ggplot(df, aes(x=eval(parse(text=feature)), fill=eval(parse(text=label_column)))) +
    geom_histogram(alpha=0.7, position="identity", aes(y = ..density.., text = ..density..), color = "gray30", bins=12) +
    geom_density(alpha=0.7) +
    labs(x=feature, y = "Density")
  plt + 
    guides(fill=guide_legend(title=label_column)) +
    theme(legend.position = "none")
}
```

## 1. Data Preprocessing

### 1.1 Load Data

```{r}
# load raw data
df.patients.raw <- read_csv("covid_individual_data.csv")
glimpse(df.patients.raw)
```

```{r}
# select variables of interest
df.patients = df.patients.raw %>%
  select(age, sex, symptoms, chronic_disease_binary, outcome)
glimpse(df.patients)
```

### 1.2 Data Quality Report

```{r}
general_data_report(df.patients)
```

```{r}
# inspect each variable
# sort(unique(df.patients$age))
# sort(unique(df.patients$sex))
# sort(unique(df.patients$symptoms))
# sort(unique(df.patients$chronic_disease_binary))
# sort(unique(df.patients$outcome))
```

### 1.3 Age

```{r}
# take the mean of age intervals
age_intervel_adjust = function(txt) {
  if(grepl("-", txt)) {
    age1 = as.numeric(strsplit(txt, "-")[[1]][1])
    age2 = as.numeric(strsplit(txt, "-")[[1]][2])
    return(as.character(mean(c(age1, age2))))
  }
  else {return(txt)}
}

# adjust ages measured in weeks/months to years
age_unit_adjust = function(txt) {
  if(grepl("month", txt)) {
    digit = as.numeric(strsplit(txt, " month")[[1]][1])
    output = round(digit/12)
    return(as.character(output))
  }else if(grepl("week", txt)) {
    digit = as.numeric(strsplit(txt, " week")[[1]][1])
    output = round(digit/52.1429)
    return(as.character(output))
  }else {return(txt)}
}

```

```{r}
# correct age intervals
df.patients$age = sapply(df.patients$age, age_intervel_adjust)
df.patients$age = sapply(df.patients$age, age_unit_adjust)

# correct ages measured in weeks/months to years
df.patients$age = round(as.numeric(df.patients$age))

# correct unreasonable age numbers (e.g. > 150) NA
df.patients$age[df.patients$age > 110 & !is.na(df.patients$age)] = NA
```

```{r}
sort(unique(df.patients$age))
```

### 1.4 Sex

```{r}
df.patients$sex = factor(df.patients$sex, levels = c("male", "female"), labels = c("Male", "Female"))

table(df.patients$sex, exclude = NULL)
```

### 1.5 Symptoms

```{r}
# remove incorrect data which contain "/"
df.patients$symptoms[sapply(df.patients$symptoms, function(x) grepl("/", x)) & !is.na(df.patients$symptoms)] = NA

# make all characters lowercases
df.patients$symptoms[!is.na(df.patients$symptoms)] = sapply(df.patients$symptoms[!is.na(df.patients$symptoms)], tolower)
```

```{r}
df.symptoms = df.patients %>%
  select(text = symptoms) %>%
  filter(!is.na(text))

df.symptoms$text = sapply(df.symptoms$text, clean_txt)
```

```{r}
# unigram
tidy.symptoms.uni = df.symptoms %>%
  unnest_tokens(word, text)

unigram_freq = tidy.symptoms.uni %>%
  count(word, sort = TRUE) %>%
  # filter(n > 2) %>%
  mutate(word = reorder(word, n))

# bigram
tidy.symptoms.bi = df.symptoms %>%
  unnest_tokens(word, text, token = "ngrams", n = 2)

bigram_freq = tidy.symptoms.bi %>%
  count(word, sort = TRUE) %>%
  # filter(n > 2) %>%
  mutate(word = reorder(word, n))


# trigram
tidy.symptoms.tri = df.symptoms %>%
  unnest_tokens(word, text, token = "ngrams", n = 3)

trigram_freq = tidy.symptoms.tri %>%
  count(word, sort = TRUE) %>%
  # filter(n > 2) %>%
  mutate(word = reorder(word, n))

# trigram
tidy.symptoms.tri = df.symptoms %>%
  unnest_tokens(word, text, token = "ngrams", n = 3)

trigram_freq = tidy.symptoms.tri %>%
  count(word, sort = TRUE) %>%
  # filter(n > 2) %>%
  mutate(word = reorder(word, n))

# quadgram
tidy.symptoms.quad = df.symptoms %>%
  unnest_tokens(word, text, token = "ngrams", n = 4)

quad_freq = tidy.symptoms.quad %>%
  count(word, sort = TRUE) %>%
  # filter(n > 2) %>%
  mutate(word = reorder(word, n))
```


```{r}
# save the ngram frequency outputs to a excel file
# write.xlsx(data.frame(unigram_freq), file="symptoms_ngram.xlsx", sheetName="unigram", row.names=FALSE)
# write.xlsx(data.frame(bigram_freq), file="symptoms_ngram.xlsx", sheetName="bigram", append=TRUE, row.names=FALSE)
# write.xlsx(data.frame(trigram_freq), file="symptoms_ngram.xlsx", sheetName="trigram", append=TRUE, row.names=FALSE)
# write.xlsx(data.frame(quad_freq), file="symptoms_ngram.xlsx", sheetName="quadgram", append=TRUE, row.names=FALSE)

```

```{r}
# create a dictionary and a new variable/column for each identified symptom
respiratory_distress_syndrome = c("respiratory distress")
respiratory_failure = c("respiratory failure")
chest_distress = c("chest distress", "chest tightness")
shortness_of_breath = c("shortness of breath", "anhelation", "dyspnea", "dyspnea")
heart_failure = c("heart failure")
joint_pain = c("joint pain")
runny_nose = c("runny nose")
septic_shock = c("septic shock")
sore_throat = c("sore throat")
aching = c("aching", "aches") 
anorexia = c("anorexia")
arrhythmia = c("arrhythmia")
backache = c("backache")
rhinitis = c("coriza", "rhinorrhea")
cough = c("cough", "coughing", "couh", "dcough")
diarrhea = c("diarrhea", "diarrheoa", "diarrhoea")
dizziness = c("dizziness")
fatigue = c("fatigue")
fever = c("fever")
flu = c("flu")
headache = c("headache")
infarction = c("infarction")
malaise = c("malaise")
myalgia = c("myalgia", "mialgia", "milagia")
nausea = c("nausea")
phlegm = c("sputum", "phlegm")
pneumonia = c("pneumonia", "pneumonitis")
sepsis = c("sepsis")
sneezing = c("sneezing")
soreness = c("soreness")
vomiting = c("vomiting")

# if the original symptom description contains any of the elements in the dictionary, then the corresponding symptom variable will be 1, otherwise, it will be 0.
generate_symptom = function(txt, dict) {
  for(item in dict) {
    if(grepl(item, txt)) {
      return(1)
    }
  }
  return(0)
}

symptoms = df.patients$symptoms
df.patients$respiratory_distress_syndrome = as.factor(sapply(symptoms, function(x) generate_symptom(x, respiratory_distress_syndrome)))
df.patients$respiratory_failure = as.factor(sapply(symptoms, function(x) generate_symptom(x, respiratory_failure)))
df.patients$chest_distress = as.factor(sapply(symptoms, function(x) generate_symptom(x, chest_distress)))
df.patients$shortness_of_breath = as.factor(sapply(symptoms, function(x) generate_symptom(x, shortness_of_breath)))
df.patients$heart_failure = as.factor(sapply(symptoms, function(x) generate_symptom(x, heart_failure)))
df.patients$joint_pain = as.factor(sapply(symptoms, function(x) generate_symptom(x, joint_pain)))
df.patients$runny_nose = as.factor(sapply(symptoms, function(x) generate_symptom(x, runny_nose)))
df.patients$septic_shock = as.factor(sapply(symptoms, function(x) generate_symptom(x, septic_shock)))
df.patients$sore_throat = as.factor(sapply(symptoms, function(x) generate_symptom(x, sore_throat)))
df.patients$aching = as.factor(sapply(symptoms, function(x) generate_symptom(x, aching)))
df.patients$anorexia = as.factor(sapply(symptoms, function(x) generate_symptom(x, anorexia)))
df.patients$arrhythmia = as.factor(sapply(symptoms, function(x) generate_symptom(x, arrhythmia)))
df.patients$backache = as.factor(sapply(symptoms, function(x) generate_symptom(x, backache)))
df.patients$rhinitis = as.factor(sapply(symptoms, function(x) generate_symptom(x, rhinitis)))
df.patients$cough = as.factor(sapply(symptoms, function(x) generate_symptom(x, cough)))
df.patients$diarrhea = as.factor(sapply(symptoms, function(x) generate_symptom(x, diarrhea)))
df.patients$dizziness = as.factor(sapply(symptoms, function(x) generate_symptom(x, dizziness)))
df.patients$fatigue = as.factor(sapply(symptoms, function(x) generate_symptom(x, fatigue)))
df.patients$fever = as.factor(sapply(symptoms, function(x) generate_symptom(x, fever)))
df.patients$flu = as.factor(sapply(symptoms, function(x) generate_symptom(x, flu)))
df.patients$headache = as.factor(sapply(symptoms, function(x) generate_symptom(x, headache)))
df.patients$infarction = as.factor(sapply(symptoms, function(x) generate_symptom(x, infarction)))
df.patients$malaise = as.factor(sapply(symptoms, function(x) generate_symptom(x, malaise)))
df.patients$myalgia = as.factor(sapply(symptoms, function(x) generate_symptom(x, myalgia)))
df.patients$nausea = as.factor(sapply(symptoms, function(x) generate_symptom(x, nausea)))
df.patients$phlegm = as.factor(sapply(symptoms, function(x) generate_symptom(x, phlegm)))
df.patients$pneumonia = as.factor(sapply(symptoms, function(x) generate_symptom(x, pneumonia)))
df.patients$sepsis = as.factor(sapply(symptoms, function(x) generate_symptom(x, sepsis)))
df.patients$sneezing = as.factor(sapply(symptoms, function(x) generate_symptom(x, sneezing)))
df.patients$soreness = as.factor(sapply(symptoms, function(x) generate_symptom(x, soreness)))
df.patients$vomiting = as.factor(sapply(symptoms, function(x) generate_symptom(x, vomiting)))
```

### 1.6 Chronic Disease Binary

```{r}
df.patients$chronic_disease_binary = as.factor(df.patients$chronic_disease_binary)
table(df.patients$chronic_disease_binary, exclude = NULL)
```

### 1.7 Outcome

```{r}
# correct all so that the outcome is either 'death' and 'Active/Recovered'
death_dict = c("dead", "Dead", "death", "Death", "Deceased", "died", "Died")
na_dict = c("mohfw.gov.in")
url_prefix = "httlp"

adjust_outcome = function(txt) {
  if(txt %in% death_dict) {return("Death")}
  else if(txt %in% na_dict | url_prefix %in% txt | is.na(txt)) {return(NA)}
  else {return("Active/Recovered")}
}

df.patients$outcome = sapply(df.patients$outcome, adjust_outcome)
df.patients$outcome = as.factor(df.patients$outcome)

table(df.patients$outcome, exclude = NULL)
```

### 1.5 Data Summary Report

```{r}
data_report_disc(df.patients[,c(2,4,5)])
data_report_cont(df.patients[,1])
```
 
## 2. Common Symptoms

### 2.1 Overall

```{r}
df.patients_with_symptoms = df.patients %>% filter(!is.na(symptoms))
df.symptoms = tibble("symptom" = c("Respiratory Distress Syndrome", "Respiratory Failure", "Chest Distress", "*Shortness of Breath",
                                   "Heart Failure", "Joint Pain", "Runny Nose", "Septic Shock", "Sore Throat", "Aching", "Anorexia",
                                   "Arrhythmia", "Backache", "Rhinitis", "Cough", "Diarrhea", "Dizziness", "Fatigue", "Fever",
                                   "Flu", "Headache", "Infarction", "Malaise", "Myalgia", "Nausea", "Phlegm", "*Pneumonia", "Sepsis",
                                   "Sneezing", "Soreness", "Vomiting"))
value = c()
for(i in 6:ncol(df.patients_with_symptoms)) {
  value = c(value, table(df.patients_with_symptoms[,i])[2])
}

df.symptoms$value = value
df.symptoms = df.symptoms %>% mutate(symptom = reorder(symptom, value))
df.symptoms
# saveRDS(df.symptoms, "df_symptoms.rds")
```

```{r}
normalize <- function(x)
{
  x = x^0.6
  return((x- min(x)) /(max(x)-min(x)))
}

ggplotly(ggplot(df.symptoms, aes(symptom, normalize(value)*10, fill = normalize(value)*10, text = paste0(symptom,"\n","Prevalence Level:", round(normalize(value)*10, 2)))) +
  geom_bar(stat = "identity", alpha = 0.85) +
  ylim(0, 10) +
  ylab("Prevalence Level") +
  xlab("Symptom") +
  theme(legend.position = "none") +
  scale_fill_gradient(low="antiquewhite", high="darkorange") + 
  coord_flip(), tooltip = c("text")) %>% layout(height = 1000, width = 800)
```

## 3. Demographic Distributions

```{r}
saveRDS(df.patients, "df_patients.rds")
```

### 3.1 Age - Overall

```{r}
ggplotly(ggplot(data = df.patients[!is.na(df.patients$age),], aes(x = "age", y = age)) +
  geom_boxplot(fill = "darkorange") +
  xlab(NULL) +
  ylab("Age") +
  coord_flip())

ggplotly(ggplot(data = df.patients %>% filter(!is.na(age)), aes(age)) + 
  geom_histogram(alpha=0.7, position="identity", aes(y = ..density..), bins = 18, fill = "darkorange", color = "gray30") +
  geom_density(alpha=0.7, color = "gray30") +
  xlab("Age") +
  ylab("Proportion")
  )
```

### 3.2 Age - Stratified by Outcome

```{r}
ggplotly(ggplot(data = df.patients %>% filter(!is.na(outcome)), aes(y = age, x = outcome, fill = outcome)) +
  geom_boxplot() + xlab("Outcome") + ylab("Age") + theme(legend.position = "none"))

ggplotly(plot_multi_histogram(df.patients %>% filter(!is.na(outcome)), feature = "age", label_column = "outcome"), tooltip = c("text"))
```

### 3.3 Sex - Overall

```{r}
ggplotly(ggplot(data = df.patients %>% filter(!is.na(sex)), aes(x = sex, fill = sex, text = round(..count../sum(..count..), 2))) +
           geom_bar(aes(y = ..count../sum(..count..))) +
           xlab("Gender") +
           ylab("Proportion") +
           theme(legend.position = "none") +
           coord_flip()
    , tooltip = c("text"))
```

### 3.4 Sex - Stratified by Outcome

```{r}
ggplotly(ggplot(data = df.patients %>% filter(!is.na(outcome) & !is.na(sex)), aes(x = outcome)) +
           geom_bar(aes(fill=sex), position="fill") +
           xlab("Gender") +
           ylab("Proportion") +
           theme(legend.position = "none")
         )
```

## 4. Association Analysis & Risk Factor Identification

Only retain observations with recorded outcomes.

```{r}
df.patients_with_outcome = df.patients %>% 
  filter(!is.na(outcome)) %>%
  select(-c(symptoms))

nrow(df.patients_with_outcome)
```

Only retain observations with non-null values for all variables.

```{r}
nrow(df.patients %>% drop_na())
```

### 4.1 2 Sample t-test for Age

```{r}
ggqqplot(df.patients_with_outcome$age[df.patients_with_outcome$outcome == "Active/Recovered"]) # seems alright
ggqqplot(df.patients_with_outcome$age[df.patients_with_outcome$outcome == "Death"]) # seems alright

var.test(age ~ outcome, data = df.patients_with_outcome) # unequal variance

t.test(age ~ outcome, data = df.patients_with_outcome, var.equal = FALSE)
```

### 4.2 Chi-Square (2x2) Tests ([Symptom, Sex] x Outcome)

#### 4.2.1 Sex

```{r}
sex.matrix = matrix(c(table(df.patients_with_outcome[df.patients_with_outcome$outcome == "Active/Recovered",]$sex),table(df.patients_with_outcome[df.patients_with_outcome$outcome == "Death",]$sex)),ncol=2)
sex.matrix

outcome = chisq.test(sex.matrix, correct = F)
outcome
```

#### 4.2.2 Symptoms

```{r}
p_values = c()
for(i in 5:ncol(df.patients_with_outcome)) {
  r1 = table(df.patients_with_outcome[df.patients_with_outcome$outcome == "Active/Recovered", i])
  r2 = table(df.patients_with_outcome[df.patients_with_outcome$outcome == "Death", i])
  m = matrix(c(r1, r2), ncol = 2)
  p_values = c(p_values, fisher.test(m)$p.value)
  print(colnames(df.patients_with_outcome)[i])
  print(m)
  print(fisher.test(m)$p.value)
}

df.symptoms.pvalues = tibble("symptom" = colnames(df.patients_with_outcome)[5:ncol(df.patients_with_outcome)],
                             "p.value" = p_values)

p_value_threshold = 0.01/nrow(df.symptoms.pvalues)
df.symptoms.pvalues %>% filter(p.value <= p_value_threshold)
```

### 4.3 Classifiation Models

#### 4.3.0 Traing/Test Split

```{r}
set.seed(8)
df = df.patients_with_outcome %>% drop_na()
rm_subset = c()
for(i in 5:35) {
  if(table(unlist(df[,i]))[2] == 0) {
    rm_subset = c(rm_subset, i)
  }
}
df = df[,-rm_subset]
df$outcome = factor(df$outcome, levels = c("Active/Recovered", "Death"), labels = c(0,1))

split = sample.split(df$outcome, SplitRatio = 0.7)
training_set = subset(df, split == TRUE)
test_set = subset(df, split == FALSE)
```

#### 4.3.1 Genreal Logistic Regression

```{r}
classifier = glm(formula = outcome ~ .,
                 family = binomial,
                 data = training_set)

prob_pred = predict(classifier, type = 'response', newdata = test_set[-4])

# model diagnostics
sum(vif(classifier) > 4)
optCutOff <- optimalCutoff(test_set$outcome, prob_pred)[1] 
misClassError(test_set$outcome, prob_pred, threshold = optCutOff)
sensitivity(test_set$outcome, prob_pred, threshold = optCutOff)
specificity(test_set$outcome, prob_pred, threshold = optCutOff)
confusionMatrix(test_set$outcome, prob_pred, threshold = optCutOff)
plotROC(test_set$outcome, prob_pred)
```

```{r}
set.seed(12)
default_glm_mod = train(
  form = outcome ~ .,
  data = df,
  trControl = trainControl(method = "cv", number = 5),
  method = "glm",
  family = "binomial"
)

default_glm_mod$results
summary(default_glm_mod)
default_glm_mod$coefnames
```

#### 4.3.2 Regularized Logistic Regression

```{r}
lasso = train(form = outcome ~ .,
             data = df,
             trControl = trainControl(method = "cv", number = 5),
             method = "glmnet",
             tuneGrid = expand.grid(alpha = 1, lambda = seq(0.00001, 2, length = 20))
             ) 
lasso$results
```

```{r}
ridge = train(form = outcome ~ .,
             data = df,
             trControl = trainControl(method = "cv", number = 5),
             method = "glmnet",
             tuneGrid = expand.grid(alpha = 0, lambda = seq(0.00001, 2, length = 20))
             ) 
ridge$results
```



